# How to Create a TravelSurevy Object

We use the 2017 US NHTS data for an example.  The data can be found
here: https://nhts.ornl.gov/assets/2016/download/Csv.zip.

```{r libs, eval = TRUE, echo = TRUE, results = "hide", warning = FALSE, error = FALSE, message = FALSE, fig.width = 8}
library("tidyverse")
library("devtools")
library("ITHIM")
```
## Household Data

First we create a data frame with the household data.  It requires two
variables _houseID_ and _location_.  _location_ can be whatever the
user wants.  Identifier variables like _houseID_ and later _subjectID_
are character class.

```{r hhpub, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
hhpub.raw <- read.csv(file = "./data/hhpub.csv", stringsAsFactors = FALSE)

hhpub <- within(hhpub.raw,{
    houseID = as.character(HOUSEID)
    location = factor(as.character(HHSTATE))
})

hhpub <- hhpub %>% select(houseID, location)
```

## Person Data

Second is the person (or individual) table.  It contains four
variables; _houseID_, _subjectID_, _age_ and _sex_.  The variables
_age_ and _sex_ are factors with levels "child", "adult", "senior" and
"M" and "F", respectively.

```{r perpub, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
perpub.raw <- read.csv(file = "./data/perpub.csv", stringsAsFactors = FALSE)

perpub <- within(perpub.raw,{
    houseID = as.character(HOUSEID)
    subjectID = as.character(PERSONID)
    sex = as.factor(ifelse(R_SEX=="1", "M", ifelse(R_SEX=="2", "F", NA)))
    age = factor(ifelse(!is.na(R_AGE), ifelse(R_AGE <= 18, "child", ifelse(R_AGE <= 65, "adult", "senior")), NA), levels = c("child","adult","senior"))
})

perpub <- perpub %>% select(houseID, subjectID, sex, age)
```

## Trip Data

Finally we have the trip data.  The _duration_ variable is numeric and
_mode_ is a factor with levels "walk", "cycle", "other".

```{r trippub, eval = TRUE, echo = TRUE, results = "show", warning = FALSE, error = FALSE, message = FALSE}
trippub.raw <- read.csv(file = "./data/trippub.csv", stringsAsFactors = FALSE)

trippub <- within(trippub.raw,{
    houseID = as.character(HOUSEID)
    subjectID = as.character(PERSONID)
    duration = ifelse(as.numeric(TRVLCMIN) < 0, as.numeric(NA), as.numeric(TRVLCMIN))
    mode = factor(ifelse(TRPTRANS %in% c(-9,-8,-7,97), NA, ifelse(TRPTRANS == 1, "walk", ifelse(TRPTRANS == 2, "cycle", "other"))), levels = c("walk","cycle","other"))
})

trippub <- trippub %>% select(houseID, subjectID, duration, mode)
```

```{r TravelSurvey, eval = TRUE, echo = TRUE, warning = FALSE, error = FALSE, message = FALSE}
NHTS <- new("TravelSurvey", person = perpub, trip = trippub, house = hhpub)
saveRDS(NHTS, file = "./R/data/NHTS.ts.rds")
```
